---
title: "Analysis_Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis_Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

# Importing Pacakge
```{r packages, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
library(ScreenR)
ggplot2::theme_set(ggplot2::theme_light())
```

# Read Data

We will use as example a Loss of Function Genetic Screening Performed on THP1
using Metforming at Day3 and Day6. First of all the data are read.
```{r read data,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
data <-  tidyr::tibble(CountTable_THP1_CONTROL_vs_MET)
colnames(data) <- 
  c("Barcode", "T0", "T48_postPURO", "Day3_Met_A", "Day3_Met_B",
    "Day3_Met_C", "Day3_DMSO_A" ,"Day3_DMSO_B","Day3_DMSO_C",
    "Day6_Met_A", "Day6_Met_B","Day6_Met_C", "Day6_DMSO_A", 
    "Day6_DMSO_B", "Day6_DMSO_C")
data <-
  data %>%
  dplyr::mutate(Barcode = as.factor(Barcode)) %>%
  dplyr::filter(Barcode != "*") 


total_Annotation <-
  Table_Annotation %>% tibble::tibble() %>%
  dplyr::mutate(Barcode = as.factor(.$Barcode))

```

# Object Creation 
The ScreenR object is created using the function **create_screenR_object()**

```{r Createe Object,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
groups <- colnames(data)[2:length(colnames(data))]
groups <- gsub('(.*)_\\w+', '\\1', groups)
groups <- factor(x = groups, levels = unique(groups))
groups

palette <-
  c("#66c2a5", "#fc8d62", rep("#8da0cb", 3), rep("#e78ac3", 3),
    rep("#a6d854", 3), rep("#ffd92f", 3))

object <- create_screenR_object(table = data,
                                annotation = total_Annotation,
                                groups = groups,
                                replicates = c(""))

```


# Start Analysis
```{r normalizzation,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
object <- normalize_data(object)
object <- compute_data_table(object)
```


## Mapped Reads
```{r plot_mapped_reads,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot <- plot_mapped_reads(object, palette) + 
  ggplot2::theme(legend.position = "none",
        axis.text.x = element_text(angle = 40, hjust = 1)) +
  ggplot2::ggtitle("Number of Mapped Reads in each sample")

plot
```


## Boxplot
```{r  distribution_mapped_reads boxplot,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot <-  distribution_mapped_reads(object, palette,
                                   alpha = 0.8,
                                   type = "boxplot") +
  ggplot2::theme(legend.position = "none",
        axis.text.x = element_text(angle = 40, hjust = 1))  

plot 
```

## Boxplot
```{r  distribution_mapped_reads density,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot <- distribution_mapped_reads(object, palette,
                                   alpha = 0.5,
                                   type = "density") +
  ggplot2::theme(legend.position = "none")  
```

## Control Genes
```{r Control Genes,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
object@data_table %>%
  dplyr::filter(Gene %in% c("RPL30", "PSMA1", "LUC")) %>%
  
  ggplot2::ggplot(., aes(x=Sample, y=Frequency, fill=Treatment)) +
  ggplot2::geom_boxplot(alpha = 0.9, outlier.shape = NA) +
  ggplot2::geom_jitter(shape=16, size = 0.7, aes(colour = Treatment)) +
  ggplot2::scale_fill_manual(values = unique(palette)) +
  ggplot2::scale_color_manual(values =  unique(palette)) +
  ggplot2::scale_alpha_manual(values=c(1, 0.1)) +
  ggplot2::theme_light()  +
  ggplot2::ylab("Normalized Mapped Reads") +
  ggplot2::theme(axis.ticks = element_line(size = 0.3),
       # axis.text.x = element_text(angle = 25),
        legend.position = "none", legend.direction = "horizontal")  +
  ggplot2::coord_flip()  +
  ggplot2::scale_x_discrete(limits = rev(unique(object@data_table$Sample))) +
  ggplot2::facet_wrap("Gene", scales = "free")
```


## Barcode Lost
```{r  plot_barcode_lost,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot <-  plot_barcode_lost(screenR_Object = object,
                           palette = palette) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "none",
        axis.text.x = element_text(angle = 40, hjust = 1))
plot
```

## Plot MDS {.tabset}

### For Sample
```{r  Plot MDS Sample,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
plot_MDS(screenR_Object = object, palette = palette)
```


### For Treatment
```{r  Plot MDS Treatment,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
GGgroups <- gsub(".*_","", groups)
pal <- c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", 
         "#66A61E", "#E6AB02", "#A6761D","#666666")
plot_MDS(screenR_Object = object, 
         palette = pal, 
         groups = factor(x = GGgroups,
                         levels = unique(GGgroups)))


```

### For Day
```{r  Plot MDS Day,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
GGgroups <- sub("_.*", "", groups)
plot_MDS(screenR_Object = object, 
         palette = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), 
         groups = factor(x = GGgroups,
                         levels = unique(GGgroups)))
```

## Compute Metrics
```{r  compute_metrics,fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
# 2DG
data_with_measure_Met <- list(
  Day3 = compute_metrics(object, control = "DMSO", 
                         treatment = "Met", day = "Day3"),
  Day6 = compute_metrics(object, control = "DMSO", 
                         treatment = "Met", day = "Day6"))


plot_Zscore_distribution(data_with_measure_Met,
                         alpha = 0.8)  +
  ggplot2::scale_fill_manual(values =  c("#1B9E77", "#D95F02")) + 
  ggplot2::theme(legend.position = "top") 
```


## Z-score hit
```{r  Z-score hit, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
zscore_hit_Met <- list(
  Day3 = find_zscore_hit(table_treate_vs_control = data_with_measure_Met$Day3,
                number_barcode = 7,
                metric = "median"),
  Day6 = find_zscore_hit(table_treate_vs_control = data_with_measure_Met$Day6,
                number_barcode = 7,
                metric = "median"))
```   


## CAMERA
```{r  CAMERA, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
groupss <- c(rep("T0/T48",4),as.character(groups[5:length(groups)]))


matrix_model <- model.matrix(~ 0 + groups)
colnames(matrix_model) <- unique(groups)

camera_hit_Met <- list(
  Day3 = find_camera_hit(screenR_Object = object,
                              matrix_model = matrix_model,
                              contrast = "Day3_Met"),
  Day6 = find_camera_hit(screenR_Object = object,
                              matrix_model = matrix_model,
                              contrast = "Day6_Met"))
```

## ROAST
```{r ROAST, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
roast_hit_Met <- list(
  Day3 = find_roast_hit(screenR_Object = object,
                              matrix_model = matrix_model,
                              contrast = "Day3_Met"),
  Day6 = find_roast_hit(screenR_Object = object,
                              matrix_model = matrix_model,
                              contrast = "Day6_Met"))
```


## Find Common Hit 
```{r Common Hit, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
common_hit_Met_at_least_2 <- list(
  Day3 = find_common_hit(zscore_hit_Met$Day3, camera_hit_Met$Day3,
                         roast_hit_Met$Day3, common_in = 2),
  Day6 = find_common_hit(zscore_hit_Met$Day6, camera_hit_Met$Day6,
                         roast_hit_Met$Day6, common_in = 2))

common_hit_Met_at_least_3  <- list(
  Day3 = find_common_hit(zscore_hit_Met$Day3, camera_hit_Met$Day3,
                         roast_hit_Met$Day3, common_in = 3),
  Day6 = find_common_hit(zscore_hit_Met$Day6, camera_hit_Met$Day6,
                         roast_hit_Met$Day6, common_in = 3))
```














<!-- ## Plot Barcode Hit -->
<!-- ```{r Barcode Hit, fig.height=7, fig.width=10, message=FALSE, warning=FALSE} -->
<!-- contrast_Day21_2DG <- makeContrasts(Day21_2DG-Day21_DMSO, levels=matrix_model) -->
<!-- plot_barcode_hit(screenR_Object = object, -->
<!--                  matrix_model = matrix_model, -->
<!--                  contrast = contrast_Day21_2DG, -->
<!--                  hit_common = common_hit_2DG_at_least_3$Day21, -->
<!--                  gene = "KDM1A") -->

<!-- plot_barcode_trend(list_data_measure = data_with_measure_2DG, -->
<!--                    genes = c("KDM1A"), -->
<!--                    n_col = 1,color = ggsci::pal_jco()(10))  -->

<!-- data_with_measure_2DG$Day21  %>% arrange(Log2FC) %>%  filter(Gene == "KDM1A") -->

<!-- library(edgeR) -->
<!-- matrix <- -->
<!--   as.matrix(data[, 2: dim(data)[2]]) -->


<!-- rownames(matrix) <- data$Barcode -->

<!-- DGEList <- DGEList(counts = matrix, -->
<!--                    group = factor(groups), genes = total_Annotation) -->
<!-- xglm <- estimateDisp(DGEList, matrix_model) -->
<!-- fit <- glmFit(xglm, matrix_model) -->

<!-- contrast_Day21_2DG <- makeContrasts(Day21_2DG-Day21_DMSO, levels=matrix_model) -->
<!-- lrt_Day21_2DG <- glmLRT(fit, contrast = contrast_Day21_2DG) -->

<!-- genesymbols <- as.character(DGEList$genes[, 1]) -->
<!-- genesymbollist2 <- list() -->
<!-- unq <- unique(genesymbols) -->
<!-- unq <- unq[!is.na(unq)] -->
<!-- for(i in unq) { -->
<!--   sel <- genesymbols == i & !is.na(genesymbols) -->
<!--   if(sum(sel)>3) -->
<!--     genesymbollist2[[i]] <- which(sel) -->
<!-- } -->


<!-- barcodeplot(statistics = lrt_Day21_2DG$table$logFC, -->
<!--             index=genesymbollist2[["KDM1A"]], -->
<!--  main= paste("Day21: Barcode plot for Gene", "KDM1A", sep = " ") , -->
<!--  labels=c("Negative logFC", "Positive logFC"), -->
<!--  quantile=c(-0.5,0.5)) -->



<!-- prova <- object@annotation_table  %>%  -->
<!--   select(Gene, Barcode,Library) -->


<!-- prova <- prova %>%  -->
<!--   mutate(Barcode = as.numeric(Barcode))  %>%  -->
<!--   group_by(Gene) %>%  -->
<!--   split(.$Gene, .$Barcode)  -->

<!-- prova <- map(.x = prova,.f = function(x){x %>% pull(Barcode)}) -->

<!-- barcodeplot(lrt_Day21_2DG$table$logFC,index=prova[["KDM1A"]], -->
<!--  main= paste("Day21: Barcode plot for Gene", "KDM1A", sep = " ") , -->
<!--  labels=c("Negative logFC", "Positive logFC"), -->
<!--  quantile=c(-0.5,0.5)) -->

<!-- ``` -->








